---
title: 'Tidy Tuesday: analyzing Animal Crossing data'
output: rmarkdown::github_document
always_allow_html: true
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

This week's Tidy Tuesday data consists data on revenue and attendance figures for theatres that are part of The Broadway League. You can take a look at the data and how it was collected on the [Tidy Tuesday GitHub repo](https://github.com/rfordatascience/tidytuesday/tree/master/data/2020/2020-04-28):

## Questions

I'm just going for a couple questions for this data:

* Trend in gross sales?
* What are the most successful Broadway shows?
* 

## Analysis

Let's start by loading our packages and data. 

```{r message = FALSE}
# Load packages
library(tidytuesdayR)
library(tidyverse)

# Load data (unable to use tt_load() this week)
tt_data <- tt_load(2020, week = 19)
user_reviews <- tt_data$user_reviews
```


### Which countries are violations most enforced in?

There are 120 countries where violations were enforced in, so let's first look at what that distribution of enforcement looks like.

```{r}
user_reviews %>% View()

```

