---
title: 'Tidy Tuesday: analyzing Broadway theater grosses'
output: rmarkdown::github_document
always_allow_html: true
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

This week's Tidy Tuesday data consists data on revenue and attendance figures for theatres that are part of The Broadway League. You can take a look at the data and how it was collected on the [Tidy Tuesday GitHub repo](https://github.com/rfordatascience/tidytuesday/tree/master/data/2020/2020-04-28):

## Questions

I'm just going for a couple questions for this data:

* Trend in gross sales?
* What are the most successful Broadway shows?
* 

## Analysis

Let's start by loading our packages and data. 

```{r message = FALSE}
# Load packages
library(tidytuesdayR)
library(tidyverse)
library(lubridate)

# Load data (unable to use tt_load() this week)
tt_data <- tt_load(2020, week = 18)
broadway_grosses <- tt_data$grosses
```


### Which countries are violations most enforced in?

There are 120 countries where violations were enforced in, so let's first look at what that distribution of enforcement looks like.

```{r}
yearly_grosses <- broadway_grosses %>% 
  distinct(week_ending, weekly_gross_overall) %>% 
  mutate(year = year(week_ending)) %>% 
  group_by(year) %>% 
  summarize(yearly_gross = sum(weekly_gross_overall))

plotly_yearly_grosses <- yearly_grosses %>% 
  filter(year < 2020) %>% 
  ggplot(aes(x = year, y = yearly_gross)) +
  geom_col(fill = "steelblue") + 
  scale_y_continuous(labels = scales::dollar) +
  theme_bw() + 
  labs(title = "",
       x = "Year",
       y = "Yearly box office gross")

plotly_yearly_grosses
```

```{r}
broadway_grosses %>% 
  mutate(year = year(week_ending)) %>% 
  group_by(year) %>% 
  summarize(yearly_avg_ticket_price = mean(avg_ticket_price))
```

