---
title: 'Tidy Tuesday: analyzing volcano eruptions'
author: "Shon Inouye"
output: rmarkdown::github_document
always_allow_html: true
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

This week's Tidy Tuesday data consists data on volcano eruptions. As someone who was born and raised on the island of Hawaii, this data is especially interesting to me. I recall a small period in my childhood where I was terrified that one of the volcanoes would erupt and create a lava flow that would engulf my house. While that imaginary scenario didn't happen to me, there was an eruption from the [Kilauea volcano in 2018](https://en.wikipedia.org/wiki/2018_lower_Puna_eruption) that caused approximately 2000 residents to evacuate, destroyed 700 houses, and left 24 people injured. That eruption made headlines around the world and resulted in studies that helped gain a deeper understanding of that volcano and others like it.

You can take a look at this week's data and how it was collected on the [Tidy Tuesday GitHub repo](https://github.com/rfordatascience/tidytuesday/tree/master/data/2020/2020-04-28):

## Questions

* 
* 

## Analysis

Let's start by loading our packages and data. 

```{r message = FALSE}
# Load packages
library(tidytuesdayR)
library(tidyverse)
library(lubridate)

# Load data (unable to use tt_load() this week)
tt_data <- tt_load(2020, week = 20)
volcano <- tt_data$volcano
eruptions <- tt_data$eruptions
events <- tt_data$events
tree_rings <- tt_data$tree_rings
sulfer <- tt_data$sulfur

```

###

```{r}
volcano %>% View()
eruptions %>% View()
events %>% View()
# tree_rings %>% View()
sulfer %>% View()
```


```{r}
eruptions_clean <- eruptions %>% 
  mutate(eruption_start = ymd(start_year, start_month, start_day))

eruptions_clean %>%   
  ggplot(aes(x = ))
```






```{r}
events %>% 
  ggplot(aes(x = eruption_start_year)) + 
  geom_histogram()

sulfer_eruptions <- eruptions %>% 
  filter(start_year >= 500,
         start_year <= 705,
         vei > 3) 

sulfer %>% 
  ggplot(aes(x = year)) + 
  geom_point(aes(y = neem), alpha = 0.3, color = "steelblue") + 
  geom_point(aes(y = wdc), alpha = 0.3, color = "tomato") + 
  geom_vline(aes(xintercept = start_year), alpha = 0.3,
             data = sulfer_eruptions)
```

